---
title: "Out of Africa Hypothesis"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
---

```{r}
chr22 = read.csv("abbgen1k.csv", stringsAsFactors = F)
```

```{r}
snpEuro <- chr22[,10:90]
```

```{r}
snpAfr <- chr22[,91:179]
```
To count the total number of ALT alleles in each SNPs
```{r}
Altsnps <- function(df1){
  maincnt <- c()
  x <- nrow(df1)
  y <- ncol(df1)
  for (i in 1:x){
    cnt = 0
    for (j in 1:y){
      if (df1[i,j] == "0|1" | df1[i,j] == "1|0"){
        cnt = cnt + 1
      }
    else if (df1[i,j] == "1|1"){
      cnt = cnt + 2
    }  
    }
    maincnt <- append(maincnt, cnt)
  }
  return (maincnt)
}
```

```{r}
colinfo <- chr22[,10:179]
xinfo <- Altsnps(colinfo)
#xinfo
```

To count the number of SNPs with greater than zero ALT frequency:
```{r}
MoreThanzero <- function(df1){
  maincnt = 0
  for (i in 1:nrow(df1)){
    cnt = 0
    for (j in 1:ncol(df1)){
      if (df1[i,j] == "1|0" | df1[i,j] == "0|1" | df1[i,j] == "1|1"){
        cnt = cnt + 1
    }
  }
  if (cnt != 0){
    maincnt = maincnt +1
  }
  }
  return (maincnt)
}
```

To count the number of SNPs in the european population with greater than zero ALT frequency
```{r}
MoreThanzero(snpEuro)
```

To count the number of SNPs in the African population with greater than zero ALT frequency
```{r}
MoreThanzero(snpAfr)
```

**Due to the out-of-Africa hypothesis, the number of SNPs with greater than zero ALT frequency in the African population is more than the European population.**


Function to get the allele freq spectrum
```{r}
allelefreq <- function(mydf, k){
  maincnt <- vector("list", k)
  maincnt <- c(1:k)*0
  for (i in 1:nrow(mydf)){
    cnt = 0
    for (j in 1:ncol(mydf)){
      if(mydf[i,j] == "1|0" | mydf[i,j] == "0|1"){
        cnt = cnt + 1
      }
      else if (mydf[i,j] == "1|1"){
        cnt = cnt + 2
      }
    }
    for (a in 1:k){
      if (cnt == a){
        maincnt[a] = maincnt[a] + 1
      }
    }
  }
  return (maincnt)
}
```

```{r}
dim(snpEuro)
```

```{r}
europlot <- allelefreq(snpEuro, 81*2)
```

```{r}
barplot(europlot)
```

```{r}
dim(snpAfr)
```

```{r}
afrplot <- allelefreq(snpAfr, 89*2)
```

```{r}
barplot(afrplot)
```

**The last bar in the african population is higher than the bar in the european population plot. This is indicative of the fact that out-of-Africa hypothesis is correct.**


To compute the bootstrap 95% confidence interval for the average pairwise diversity statistic between european and African population separately:
```{r}
ntavg <- function(df1){
 x1 <- c(1,3)
 n = sample(x1,1)
 m = sample(x1,1)
 rand_euro <- df1[sample(ncol(df1), size = 2)]
 x <- substring(rand_euro[,1],n,n)
 y <- substring(rand_euro[,2],m,m)
 return(mean(x != y))
 }
```

```{r}
ntavg(snpEuro)
ntavg(snpAfr)
```

```{r}
avgSum <- function(df1){
 fiftyavgSum <- mean(replicate(50, ntavg(df1)))
 return (fiftyavgSum)
}
```

```{r}
avgSum(snpEuro)
avgSum(snpAfr)
```

```{r}
finalfunc <- function(df1){
 mylist <- c() 
 hunAvg <- replicate(100, avgSum(df1))
 sortedAvg <- sort(hunAvg)
 thirdval <- sortedAvg[3]
 ninsevenVal <- sortedAvg[97]
 mylist <- c(thirdval, ninsevenVal)
 return (mylist)
}
```

```{r}
finalfunc(snpEuro)
finalfunc(snpAfr)
```

**The average pairwise diversity within Europe is lesser than both within African population and between European and African population.**


To compute the bootstrap 95% confidence interval for the average pairwise diversity statistic between european and African population:
```{r}
ntavgmix <- function(dfeuro, dfafr){
 x1 <- c(1,3)
 n = sample(x1,1)
 m = sample(x1,1)
 rand_euro <- dfeuro[sample(ncol(dfeuro), size = 1)]
 rand_afr <- dfafr[sample(ncol(dfafr), size = 1)]
 x <- substring(rand_euro[,1],n,n)
 y <- substring(rand_afr[,1],m,m)
 return(mean(x != y))
 }
```

```{r}
ntavgmix(snpEuro, snpAfr)
```

```{r}
avgSumMix <- function(dfeuro, dfafr){
 fiftyavgSum <- mean(replicate(50, ntavgmix(dfeuro, dfafr)))
 return (fiftyavgSum)
}
```

```{r}
avgSumMix(snpEuro, snpAfr)
```

```{r}
finalfunc2 <- function(dfeuro, dfafr){
 mylist <- c() 
 hunAvg <- replicate(100, avgSumMix(dfeuro, dfafr))
 sortedAvg <- sort(hunAvg)
 sortedAvg
 thirdval <- sortedAvg[3]
 ninsevenVal <- sortedAvg[97]
 mylist <- c(thirdval, ninsevenVal)
 return (mylist)
}
```

```{r}
finalfunc2(snpEuro, snpAfr)
```

